---
title: "Jamaica Travel Data for 2019"
author: "Dominique McIntosh"
date: '2022-07-04'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Visit Jamaica

### A dive into visitor data regarding travel to Jamaica for the year 2016:

-   Who was coming?

-   Why?

-   How many? Etc.

For more analysis & visualizations including an interactive Excel dashboard click here.

For a more detailed look at SQL queries run to generate some of the csv files, head over to Github

## Why this project?

Just sheer curiosity about the information that was available. My history running a bed & breakfast for the past few years, you start to notice certain trends about the visitors. While I had a few obvious confirmations, it was still surprisingly, a bit of an eye-opener after analyzing the raw information.

## Summary of The Data Source

The source of the travel data came from [Jamaica Open Data](https://data.gov.jm/), a public data repository provided by the Jamaican government from the various ministries that includes a variety of data from traffic to health. The most recent travel data available on this platform was for the year 2016.

-   Actual data was collected from embarkation/disembarkation cards (immigration forms) all air passengers are required to fill out on arrival.

-   N.B.: Jamaican nationals who reside overseas are included in the numbers.

-   Cruise passengers flying into the island to meet the cruise ship are not included in tourist arrival figures unless they spend at least one night in a land-based accommodation.

-   The file is a CSV file.

## Summary of Data Cleaning

-   The CSV file was imported into Excel.

-   Adjusted & formatted data:

    -   changed header titles for clarification

    -   confirmed correct data types

    -   remove some abbreviations & replaced with full word

-   Removed Duplicated (47)

-   Removed spaces (trim())

-   Corrected misspellings

-   Performed preliminary analysis: complete & accurate

## A little Analysis & Data Visualizations Using R:

#### Installing & loading the following packages & libraries:

-   tidyverse

-   dplyr

-   ggplot2

-   ggthemes

```{r, echo=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggthemes)
```

#### Importing csv files created with SQL

```{r, echo=FALSE}
travel_reason <- read.csv("reason_for_travel-SQL.csv")
airport_arrivals <- read.csv("total_travelers-port_of_entry-SQL.csv")
country <- read.csv("visitors_country-SQL.csv")
```

### Total Arrivals by Airport

```{r, echo=FALSE}
airport_arrivals %>% 
  ggplot(aes(x=port_of_entry, y=total_visitors, fill = port_of_entry, width = .8))+
  geom_bar(stat = "identity")+
  theme_clean()+ #TRYING ONE OF THE THEMES
    scale_fill_discrete(labels = c("NMIA - Kingston", 
                                 "SIA - Montego Bay"))+
  scale_y_continuous(labels = function(x) format(x, big.mark = ",",
                                                 scientific = FALSE))+ # ADDS "," to big numbers on AXES
  labs(title="Total Arrivals by Airport", subtitle = " Data for 2016", y="Total Visitors",
       x = "Port Of Entry", fill =" Airport")+
    annotate("text",x=1,label="700,000",y=600000,size=4,fontface="italic")+
  annotate("text", x=2, label="1.6 Million", y=1500000, size=4, fontface="bold.italic")
```

### Top 5 Countries that visited Jamaica

```{r, echo=FALSE}
country %>% # Piping the Data
  filter(country != "VAC") %>% # FILTERING OUT THIS ERROR
  slice_max(total_visitors, n=5) %>%  # ONLY LOAD THE TOP 5
  mutate(country = fct_reorder(country, total_visitors)) %>%  # REORDERING BY VALUE 
  ggplot(aes(x=country, y=total_visitors, fill = country, width = .9))+
  geom_bar(stat = "identity")+
  labs(title="Top 5 Countries that Visited Jamaica", subtitle = "Data for 2016", y="Total Visitors")+
  annotate("text",x=1,label="13,500",y=100000,size=4,fontface="italic")+
  annotate("text", x=2, label="22,000", y=100000, size=4, fontface="italic")+
  annotate("text", x=3, label="214,000", y=100000, size=4, fontface="italic")+
  annotate("text", x=4, label="407,000", y=100000, size=4, fontface="italic")+
  annotate("text", x=5, label="1.4 Million", y=1000000, size=4, fontface="bold.italic")+
  theme(legend.position = "none",
        axis.text.y=element_blank(), # REMOVE Y-AXIS data (too far apart)
        axis.title.x=element_blank(),
        axis.ticks.x=element_blank()# REMOVE DEFAULT X-AXIS Title - NOT NEEDED
        )

```

### Top 5 Reasons for Travel to Jamaica from The USA

```{r, echo=FALSE}
travel_reason %>% 
  filter(country == "U.S.A.") %>% 
  slice_max(total_visitors, n=5) %>% 
  mutate(reason_for_travel = fct_reorder(reason_for_travel, total_visitors)) %>% 
  ggplot(aes(x=reason_for_travel, y=total_visitors, fill = reason_for_travel, width = .8))+
  geom_bar(stat = "identity")+
  scale_y_continuous(labels = function(x) format(x, big.mark = ",",
                                                 scientific = FALSE))+ # ADDS "," to big numbers on AXES
  labs(title="Top 5 Reasons for Travel to \nJamaica from The USA",  subtitle = "Data for 2016",
       y="Total Visitors"  )+
  annotate("text",x=1,label="43,220",y=125000,size=4,fontface="bold.italic")+ #ADDING SPECIFIC NUMBERS TO Bars
  annotate("text", x=2, label="79,767", y=150000, size=4, fontface="bold.italic")+
  annotate("text", x=3, label="115,603", y=200000, size=4, fontface="bold.italic")+
  annotate("text", x=4, label="186,624", y=260000, size=4, fontface="bold.italic")+
  annotate("text", x=5, label="912,314", y=600000, size=5, fontface="bold.italic")+
  theme(legend.position = "none", # REMOVE THE LEGEND - Not Needed
        title = element_text(family = "special", face="bold"),
        axis.title.y=element_blank(),
        axis.text.x=element_text(face="bold", family="special"),
        axis.text.y = element_text(face="bold", family="special")
        )+                       
  coord_flip()
```

### Top 5 Reasons for Travel to Jamaica from Canada

```{r, echo=FALSE}
travel_reason %>% 
  filter(country == "CANADA") %>% #APPLY FILTER FOR CANADA
  slice_max(total_visitors, n=6) %>% #FILTER TO SHOW TOP 5
    mutate(reason_for_travel = fct_reorder(reason_for_travel, desc(total_visitors))) %>% 
  ggplot(aes(x= reason_for_travel, y=total_visitors)) +
  #----DECORATING THE STEM 
  geom_segment(aes(x=reason_for_travel, xend=reason_for_travel, y=0, yend=total_visitors),
               size=1,
               color="tomato")+ 
  scale_y_continuous(labels = function(x) format(x, big.mark = ",",
                                                 scientific = FALSE))+
  #DECORATING THE LOLLIPOP
  geom_point(stat="identity", size=4, 
             color="purple",
             fill=alpha("pink", .3), #---color inside circle
             alpha=1,
             shape=21,
             stroke=1) + 
  #TRYING ONE OF THE THEMES
  theme_clean()+ 
  labs(title="Top 5 Reasons for Travel to Jamaica from Canada",  subtitle = "Data for 2016",
       y="Total Visitors"  )+
  geom_text(aes(label = format(total_visitors, big.mark = ",")), vjust=-1)+ #ADDING "," to labels
  theme(legend.position = "none", # REMOVE THE LEGEND - Not Needed
        title = element_text(family = "special", face="bold"),
        axis.title.y=element_blank(),
        axis.text.x=element_text(face="bold", family="special"),
        axis.text.y = element_text(face="bold", family="special"))+                       
  coord_flip()
```

### Top 5 Reasons Travel To Jamaica from the UK

```{r, echo=FALSE}
travel_reason %>% # Piping the Data
  filter(country=="UNITED KINGDOM") %>% #Filter for the UK
  slice_max(total_visitors, n=6) %>% 
  mutate(reason_for_travel = fct_reorder(reason_for_travel, total_visitors)) %>% 
  ggplot(aes(x=reason_for_travel, y=total_visitors, fill = reason_for_travel, width = .9))+
  geom_bar(stat = "identity")+
  geom_text(aes(label = format(total_visitors, big.mark = ",")), vjust = 1.5, colour = "black")+
  scale_y_continuous(labels = function(x) format(x, big.mark = ",",
                                                 scientific = FALSE))+
  labs(title="Overall Reason for Travel to Jamaica in 2016", y="Total Visitors")+
  theme(legend.position = "none", # REMOVE THE LEGEND - Not Needed
        title = element_text(family = "special", face="bold"),
        axis.title.y=element_blank(),
        axis.text.x=element_text(face="bold", family="special"),
        axis.text.y = element_text(face="bold", family="special")
  )+                       
  coord_flip()

```

### Other than the Big Three (US,CA,UK) - who else came? Top 25

```{r, echo=FALSE}
country %>% 
  filter(country != "VAC" & country != "U.S.A." & country != "CANADA"
         & country != "UNITED KINGDOM") %>% #APPLY FILTER FOR The BIG 3
  slice_max(total_visitors, n=25) %>%
  mutate(country = fct_reorder(country, total_visitors)) %>% 
  ggplot(aes(x= country, y=total_visitors)) +
  
  scale_y_continuous(labels = function(x) format(x, big.mark = ",",
                                                 scientific = FALSE))+
  theme_clean()+
  #----DECORATING THE STEM 
  geom_segment(aes(x=country, xend=country, y=0, yend=total_visitors),
               size=1,
               color="pink")+ 
  #DECORATING THE LOLLIPOP
  geom_point(stat="identity", size=2, 
             color="purple",
             fill=alpha("purple", .3), #---color inside circle
             alpha=1,
             shape=21,
             stroke=1) + 
  labs(title="Top 25 Countries outside of 'The Big Three'",  subtitle = "Data for 2016",
       y="Total Visitors"  )+
    theme(legend.position = "none",
        title = element_text(family = "special", face="bold"), #OVERALL TITLE
        axis.title.x=element_blank(),
        axis.text.x=element_text(face="bold")#----REMOVING THE DEFAULT X-AXIS TITLES
            )+
  coord_flip()
```

### Conclusions:

-   No surprise that the 'Big Three' (US, Canada, UK) dominate the visitor numbers, however the United States far eclipsed even the other two countries by more than double their amounts!

-   The Cayman Islands travel numbers were also surprising. With a population of only 67,000, to be Jamaica's number 4 in total visitors I don't think anyone could have guessed.

-   IT'S A VIBE 🥳! VACATION is the number 1 reason people come to Jamaica (no real surprise there).

### Recommendations:

-   Jamaica's tourism product is very precious & helps to drive the country's economy. It would be beneficial for the government to encourage travel to the island targeting other markets/countries outside of the Big Three.

-   In addition, it might also be beneficial to encourage travel to the island for reasons other than vacationing, such as for business & sport (which the country dominates globally when it comes to track & field).

Well that's my little project about visitor data to Jamaica. I hope it was interesting and eye-opening!

Thanks for having a look!

Dominique
